---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Riccardo Fusaroli"
date: "October 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3 - Part 2 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia. We then looked at whether we could replicate results from the previous literature.
We now want to know whether we can automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.
Again, remember that the dataset containst 7 studies and 3 languages. Feel free to only include Danish (Study 1-4) if you feel that adds too much complexity.


Issues to be discussed your report:
- Should you run the analysis on all languages/studies at the same time? 
- Choose your best acoustic feature from part 1. How well can you diagnose schizophrenia just using it?
- Identify the best combination of acoustic features to diagnose schizophrenia using logistic regression.
- Discuss the "classification" process: which methods are you using? Which confounds should you be aware of? What are the strength and limitation of the analysis?
- Bonus question: Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them. 
- Bonus Bonus question: It is possible combine the output of multiple  classification models to improve classification accuracy. For inspiration see,
https://machinelearningmastery.com/machine-learning-ensembles-with-r/
 The interested reader might also want to look up 'The BigChaos Solution to the Netflix Grand Prize'

## Learning objectives
- Learn the basics of classification in a machine learning framework
- Design, fit and report logistic regressions
- Apply feature selection techniques

### Let's start

load packages and data

```{r}
library(pacman)
p_load(tidyverse, tidymodels, lme4, boot, caret, ranger, cvms, groupdata2, doParallel, e1071)

df <- read_csv('SchizophreniaData.csv')

df <- df %>% mutate(Diagnosis = factor(Diagnosis),
                    Study = factor(Study),
                    Subject = factor(Subject),
                    Trial = factor(Trial),
                    ID = factor(ID),
                    Gender = factor(Gender),
                    Age = factor(Age),
                    speechrate = speechrate..nsyll.dur.
                    ) %>% 
  select(- Language, - time, - f0, - speechrate..nsyll.dur.) %>%
  select(!starts_with('time'))
  select(ID, ID2, Subject, Study, Trial, Gender, Age, everything())


df <- df %>% mutate(prop_spoken_time = phonationtime..s. / dur..s.,
                    pause_duration = (dur..s. - phonationtime..s.) / npause,)

df$pause_duration <-  ifelse(df$npause == 0, 0, df$pause_duration)


# in case we need diagnosis as dummy code
df$Diagnosis1 <-  ifelse(df$Diagnosis == 'Control', 0, 1) %>% as.factor()

```


We first want to build a logistic regression to see whether you can diagnose schizophrenia from your best acoustic feature. Let's use the full dataset and calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve). You need to think carefully as to how we should (or not) use study and subject ID.

```{r}
#explore data and visualize

df %>% ggplot(aes(Diagnosis, pause_duration)) + geom_point() 


```

apply feature selection techniques?

```{r}
# find the one best feature from part 1

m1 <- glmer(Diagnosis ~ speechrate + (1|ID), family = 'binomial', df)
summary(m1)

pred_probs <- predict(m1, type = 'response', re.form = NA)
actual_value <- df$Diagnosis

d <- tibble(pred_probs, actual_value)

d$pred_category <- ifelse(d$pred_probs > 0.5, 'Schizophrenia', 'Control') %>%  as.factor()

cm <- confusionMatrix(d$pred_category, d$actual_value, positive = 'Schizophrenia')
cm

# find the performance measures 
#see output of cm

```

find the best combination of features

to find the best model, we now split into training and testing data.

```{r}

p1 <- partition(df, p = 0.2, cat_col = 'Diagnosis1', id_col = 'ID')

df_test <- p1[[1]]
df_train <- p1[[2]]

```


Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures. Alternatively, the groupdata2 and cvms package created by Ludvig are an easy solution. 

```{r}
# scale and normalize? data
df_train <- df_train %>% mutate_if(is.numeric, scale)
df_train <- df_train %>% select(ID, Diagnosis1, speechrate, pause_duration, pitch_IQR, prop_spoken_time)

#exploratory model creation and feature selection


#remove correlations
a <-cor(df_train %>%  select_if(is.numeric))
a <- as.data.frame(a)

```


```{r}
#from cvms 

clf_svm_model_fn <- function(train_data, formula, hyperparameters) {

  # Optional hyperparameters:
  #  - kernel
  #  - cost

  # Update missing hyperparameters with default values
  hyperparameters <- update_hyperparameters(
    kernel = "radial",
    cost = 1,
    hyperparameters = hyperparameters
  )

  e1071::svm(
    formula = formula,
    data = train_data,
    kernel = hyperparameters[["kernel"]],
    cost = hyperparameters[["cost"]],
    type = "C-classification",
    probability = TRUE  # Must enable probability here
  )
}

# Try the model function
m1 <- clf_svm_model_fn(train_data = df_train, formula = Diagnosis1 ~ speechrate + pitch_IQR, 
                       hyperparameters = list("kernel" = "linear"))

```


```{r}
# Predict function for binomial SVM
bnml_svm_predict_fn <- function(test_data, model, formula, hyperparameters, train_data) {
  # Predict test set
  predictions <- predict(
    object = model,
    newdata = test_data,
    allow.new.levels = TRUE,
    probability = TRUE
  )
  
  # Extract probabilities
  probabilities <- dplyr::as_tibble(attr(predictions, "probabilities"))
  
  # Return second column
  probabilities[[2]]
}

p1 <- bnml_svm_predict_fn(test_data = df_test, model = m1)
p1


```

```{r}
#create folds

set.seed(5)
df_train <- fold(df_train, 
                 k = 10, 
                 cat_col = 'Diagnosis1', 
                 id_col = 'ID', 
                 num_fold_cols = 10)

#now we can crossvalidate

doParallel::registerDoParallel()
cv_4 <- cross_validate_fn(
  data = df_train,
  formulas = "Diagnosis1 ~ speechrate + pitch_IQR",
  type = "binomial",
  model_fn = clf_svm_model_fn,
  predict_fn = bnml_svm_predict_fn,
  hyperparameters = list(
    "kernel" = c("linear", "radial"),
    "cost" = c(1, 5, 10)
  ),
  fold_cols = paste0(".folds_", 1:10)
)

# creating 1200 models, takes time

cv_4
```



```{r}
cv_4 %>% 
  mutate(`Model ID` = 1:nrow(cv_4)) %>% 
  arrange(desc(`Balanced Accuracy`)) %>% 
  select_definitions(additional_includes = c("Balanced Accuracy", "F1", "MCC", "Model ID"))

# this doesn't work

cv_4 %>% 
  # Create Model ID with values 1:8
  dplyr::mutate(`Model ID` = 1:nrow(cv_4)) %>% 
  # Order by RMSE
  dplyr::arrange(RMSE) %>% 
  # Extract formulas and hyperparameters
  select_definitions(additional_includes = c("RMSE", "Model ID"))

p <- cv_4$Predictions
p1 <- p[[1]]
p1_fold1 <- p1 %>% filter(`Fold Column` == '.folds_1')
prediction_1 <- p1_fold1$Prediction

df_test<- cbind(df_train, prediction_1)

```

```{r}
#final run?

final_svm_model <- svm(Diagnosis1 ~ speechrate + pitch_IQR, type = 'C-classification', kernel = 'linear', cost = 1, df_train)

pred <- predict(final_svm_model, df_test)


?predict

actual_value_t <- df_test$Diagnosis1

d <- tibble(pred)


```



```{r}
ev <- evaluate(
  data = df_test,
  target_col = "Diagnosis1",
  prediction_cols = "probability",
  id_col = "ID",
  id_method = "mean",
  type = "binomial"
)

most_challenging(
  data = data,
  type = "gaussian",
  obs_id_col = "participant",
  target_col = "score",
  prediction_cols = "predicted score",
  threshold = 0.20
)


```


N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.


N.N.N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?
N.N.N.N.B. A more advanced solution could rely on the tidymodels set of packages (warning: Time-consuming to learn as the documentation is sparse, but totally worth it)


